Reliable error estimation for Sobol' indices\thanks{This work is supported by the CITiES project funded by the Agence Nationale de la Recherche (grant ANR-12-MONU-0020) and by the United States National Science Foundation (grant DMS-1522687).}

### 2016-09-28 article/article_for_submission.tex
Last changes for article

As for the g-function, Tables \ref{res.bratley.Aa} and \ref{res.bratley.Ab} show averaged error bounds and averaged total number of evaluations for Variant $A.a$ and Variant $A.b$.
\begin{table}[!ht]
\caption{Averaged error bounds $\varepsilon_{\underline{S}_u}$, $\varepsilon_{\overline{S}_u}$ and total number of evaluations of Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\varepsilon_{\underline{S}_u}$ & ${err}_{\varepsilon_{\underline{S}_u}}$ & $\overline{S}_u$ & $\varepsilon_{\overline{S}_u}$ &$err_{\varepsilon_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0.$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0004$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0058$ & $0.71$ & $0.0764$ & $0.0012$ & $0.02$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0022$ & $0.12$ & $0.0343$ & $0.0020$ & $0.07$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0024$ & $0.11$ & $0.0062$ & $0.0014$ & $0$ \\ \hline
 $x_6$ & $0.0015$ & $0.0021$ & $0.06$  & $0.0062$ & $0.0010$ & $0.03$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $70 \ 129$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged error bounds $\varepsilon_{\underline{S}_u}$, $\varepsilon_{\overline{S}_u}$ and total number of evaluations of Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\varepsilon_{\underline{S}_u}$ & ${err}_{\varepsilon_{\underline{S}_u}}$ & $\overline{S}_u$ & $\varepsilon_{\overline{S}_u}$ &$err_{\varepsilon_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0005$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0019$ & $0.04$ & $0.0764$ & $0.0013$ & $0.01$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0016$ & $0.01$ & $0.0343$ & $0.0017$ & $0.02$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0003$ & $0$ & $0.0062$ & $0.0015$ & $0.04$ \\ \hline
 $x_6$ & $0.0015$ &  $0.0004$ & $0$  & $0.0062$ & $0.0011$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $68 \ 045$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Ab}
\end{table}
Variant $A.b$ provides better error bounds than Variant $A.a$ on the four small first-indices $\underline{S}_3,\underline{S}_4,\underline{S}_5$, $\underline{S}_6$ which highlights the performance of ``Correlation 2''. The discrepancy is particularly notable for input $x_3$ where Variant $A.a$ reaches a failure rate of $71\%$ against only $4\%$ for Variant $A.b$. Furthermore, Variant $A.b$ requires less evaluations than Variant $A.a$.
\begin{figure}[!ht]
\caption{Boxplots of error bounds $\varepsilon_{\underline{S}_u}$ for the four inputs $x_3,x_4,x_5,x_6$ obtained with both Variant $A.a$ and Variant $A.b$. The horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.}
\centering
\includegraphics[width=0.45\textwidth]{Images/boxplots_bis.pdf}
\label{boxplots.bratley}
\end{figure}
The boxplots represented in Figure \ref{boxplots.bratley} emphasize the latter observations and remarks.

@@

From the results on these two functions test, the main conclusion is that Variant $A.b$ performs the best. This highlights the efficiency of ``Correlation 2'' to estimate small first-order indices. Even if ``Correlation 2'' estimator originally requires more samples, this drawback is completely overridden when this estimator is included into a sequential procedure such as ours.

@@

The sequential estimation procedure proposed in this paper offers a practical solution with the construction of an error bound for the Sobol' indices. The number of points is progressively augmented until the error bound becomes lower than a user specified tolerance. The procedure presented combines Sobol' sequences with either Saltelli's strategy to estimate both first-order and total effect indices, or the replication procedure to estimate only first-order indices. Furthermore, we investigated the use of a recent estimator well-suited to the estimation of small first-order indices. The efficiency of this estimator was assessed and highlighted on two test functions and a real case model. More generally, our estimation procedure gave good results in term of precision of the error bounds with low failure rates observed across the indices.

>>>>

As for the g-function, Tables \ref{res.bratley.Aa} and \ref{res.bratley.Ab} show averaged error bounds and averaged total number of evaluations of Variant $A.a$ and Variant $A.b$.
\begin{table}[!ht]
\caption{Averaged error bounds $\varepsilon_{\underline{S}_u}$, $\varepsilon_{\overline{S}_u}$ and total number of evaluations of Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\varepsilon_{\underline{S}_u}$ & ${err}_{\varepsilon_{\underline{S}_u}}$ & $\overline{S}_u$ & $\varepsilon_{\overline{S}_u}$ &$err_{\varepsilon_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0.$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0004$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0058$ & $0.71$ & $0.0764$ & $0.0012$ & $0.02$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0022$ & $0.12$ & $0.0343$ & $0.0020$ & $0.07$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0024$ & $0.11$ & $0.0062$ & $0.0014$ & $0$ \\ \hline
 $x_6$ & $0.0015$ & $0.0021$ & $0.06$  & $0.0062$ & $0.0010$ & $0.03$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $70 \ 129$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged error bounds $\varepsilon_{\underline{S}_u}$, $\varepsilon_{\overline{S}_u}$ and total number of evaluations of Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\varepsilon_{\underline{S}_u}$ & ${err}_{\varepsilon_{\underline{S}_u}}$ & $\overline{S}_u$ & $\varepsilon_{\overline{S}_u}$ &$err_{\varepsilon_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0005$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0019$ & $0.04$ & $0.0764$ & $0.0013$ & $0.01$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0016$ & $0.01$ & $0.0343$ & $0.0017$ & $0.02$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0003$ & $0$ & $0.0062$ & $0.0015$ & $0.04$ \\ \hline
 $x_6$ & $0.0015$ &  $0.0004$ & $0$  & $0.0062$ & $0.0011$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $68 \ 045$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Ab}
\end{table}
Variant $A.b$ provides better error bounds than Variant $A.a$ on the four small first-indices $\underline{S}_3,\underline{S}_4,\underline{S}_5$, $\underline{S}_6$ which highlights the performance of ``Correlation 2''. The discrepancy is particularly notable for input $x_3$ where Variant $A.a$ reaches a failure rate of $71\%$ against only $4\%$ for Variant $A.b$. Furthermore, Variant $A.b$ requires less evaluations than Variant $A.a$.
\begin{figure}[!ht]
\caption{Boxplots of error bounds $\varepsilon_{\underline{S}_u}$ for the four inputs $x_3,x_4,x_5,x_6$ obtained with both Variant $A.a$ and Variant $A.b$. The horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.}
\centering
\includegraphics[width=0.45\textwidth]{Images/boxplots_bis.pdf}
\label{boxplots.bratley}
\end{figure}
The boxplots represented in Figure \ref{boxplots.bratley} emphasize the latter observations and remarks.

@@

From the results on these two functions test, the main conclusion is that Variant $A.b$ performs the best. This highlights the efficiency of ``Correlation 2'' to estimate small first-order indices. Even if ``Correlation 2'' originally requires more samples, this drawback is completely overridden when this estimator is included into a sequential procedure such as ours.

@@

The sequential estimation procedure proposed in this paper offers a practical solution with the construction of an error bound for the Sobol' indices. The number of points is progressively augmented until the error bound becomes lower than a user specified tolerance. The procedure presented combines Sobol' sequences with either Saltelli's strategy to estimate both first-order and total effect indices, or the replication procedure to estimate only first-order indices. Furthermore, we investigated the use of a recent estimator well-suited to the estimation of small first-order indices. The efficiency of this estimator, called ``Correlation 2'', was assessed and highlighted on two test functions and a real case model. The variant combining Saltelli's strategy and ``Correlation 2'' gave the best results in term of precision of the error bounds with low failure rates observed across the indices.
